Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-15 qpip0.tex
Add action items on QPIP0 section

In this section, we combine \cref{QPIP1thm} with results from \cite{FOCS:Mahadev18a} to create a delegation protocol for $\SampBQP$ for fully classical clients. To boost the soundness, we need  polynomial rounds, so we use results from \cite{parallelrep} to run polynomially many copies in parallel and get a constant-round protocol.

@@

\subsection{Mahadev's classical "commitment" of $XZ$ measurement}

@@

\begin{protocol}{Mahadev's 4 round commitment protocol $\PiMeasure$}
\label{proto:urmila4}
\Ethan{Still rooms for improvements, but it's not as bad as before}

Inputs:
\begin{itemize}
	\item Common input: Security parameter $1^\lambda$ where $\lambda\in\bbN$.
	\item Prover's input: a state $\rho\in\cB^{\otimes n}$ for the verifier to measure.
	\item Verifier's inputs:
		\begin{itemize}
			\item a string $h \in \{0,1\}^n$ that specifies the measurement he wants to make: if $h_i=0$, he wants a $X$ measurement on $i$-th qubit. If $h_i=1$, he wants a $Z$ measurement on $i$-th qubit.  
			\item a challenge bit $c\in\zo$. The prover needs to do a ``Testing round" if $c=0$. The prover needs to do a ``Hadamard round" if $c=1$.
		\end{itemize}
\end{itemize}

The protocol:
\begin{enumerate}
    \item The verifier generates a public and secret key pair $(pk, sk)$ on inputs $h$ and $1^\lambda$. It sends $pk$ to the prover.
    \item The prover generates a classical ``commitment" $y$ and internal state $\sigma$ on inputs $pk$ and $\rho$. He send $y$ to the verifier.
	\item The verifier sends $c$ to the prover.
    \item The prover generates a classical string $a$ on inputs $c$ and $\sigma$, and sends it to back to the verifier.
    \item If $c=0$, the verifier decides whether he is going to reject on inputs $(pk,a,y)$.
		If $c=1$, the verifier computes an $n$-bit string $z=V_f(sk,h,y,a)$ on inputs $(sk,h,y,a)$.
\end{enumerate}
\end{protocol}

\Ethan{The following modifications are mine}
Two modification have been made from the original protocol.
\begin{itemize}
	\item On the protocol's third message. Instead of randomly picking whether to run what's called a ``testing round" or a ``Hadamard round", we let the verifier decide as one of its inputs.
	\item On the protocol's fourth message, the prover can choose to abort by sending $a=\bot$. This is useless for the cheating prover but useful for proving one of our theorems.
\end{itemize}

\begin{definition}
With loss of generality,  we can assume that the testing round verification is deterministic. We denote the set of strings $a$ \Ethan{Font makes this $a$ look screwy} accepted by the verifier in the testing round with $(pk,y)$ as $\Acc_{pk,y}$.
\end{definition}

\begin{lem}[\hannote{ref}]~\label{lem:trivial-4-round-strategy}
    For any input, there exist a prover strategy $(U_{trivial,0},U_{trivial})$ for Protocol~\ref{proto:urmila4} that is accepted with probability $1-\negl{n}$ in the testing.
\end{lem}

@@

\begin{lemma}\label{lem:urmila-binding}[binding property of  Protocol~\ref{proto:urmila4}\cite{FOCS:Mahadev18a}]
If a prover's strategy in Protocol~\ref{proto:urmila4}\hannote{need clarification} results in passing the testing round with probability $1-\negl(n)$, then there must exist a $\rho$ such that for all $h$, the verifier will produce a $z$ on the Hadamard round that is $\negl(n)$-Q-computationally indistinguishable\footnote{Recall that quantum computational indistinguishability is defined in Definition~\ref{def:comp-indis}.} from $M_{XZ}(\rho,h)$\hannote{add notation}, unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

@@

By combining the commitment Protocol, $\PiMeasure$, and our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, we get a naive $\QPIP_0$ Protocol for $\SampBQP$ that has no soundness guarantee.
\Ethan{The following discussion is mine}
In particular, we combine them by replacing the verifier's measurements with a call to $\PiMeasure$. It is possible because $\VSamp$ applies only $X$ and $Z$ measurements.
Furthermore, its measurements are not adaptive, meaning measurement results do not change later measurement basis choices.

\begin{protocol}{naive $\QPIP_0$ protocol $\PiNaive$ for the $\SampBQP$ instance $(D_x)_{x\in\set{0, 1}^*}$}
	\label{proto:qpip0_naive}

	Inputs:
	\begin{itemize}
		\item Security parameter $1^\lambda$ where $\lambda\in\bbN$
		\item Error parameter $\varepsilon\in(0, 1)$
		\item Classical input $x\in\zo^n$ to the $\SampBQP$ instance
		\item Verifier's input: challenge bit $c\in\zo$ as in $\PiMeasure$
	\end{itemize}

@@

	Protocol:
	\begin{enumerate}
		\item The verifier choose a $XZ$-measurement $h$ from the distribution emergent from by Step~\ref{step:qpip1-verify} of $\PiSamp$. \Ethan{I didn't change the wording here}
			It samples $c\xleftarrow{\$}\zo$ uniformly at random.
		\item The prover prepares $\sigma$ by running Step~\ref{step:qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive} The verifier and prover run $(\PMeasure(\sigma), \VMeasure(h))(1^\lambda, 1^{1/\varepsilon}, x)$.
			If $c=1$, the verifier obtains $z_{commit}$.
		\item If $c=1$, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by using $z_{commit}$ as the measurement result he would get and outputs $(d,z_{naive})$.
	\end{enumerate}
\end{protocol}

@@

\begin{equation} \label{eq:z=verM}
    \norm{(d,z_{naive})-(Ver( M_{XZ}(\rho,h))+p_{abort}(\Rej,\bot))}_q \leq \negl(n).
\end{equation}
\hannote{actually true with subnormalize shit?}
 By Theorem~\ref{QPIP1thm}, for all $\rho$, 
 \begin{equation}
     \norm{Ver(M_{XZ}(\rho,h))-(d,z_{ideal})}\leq O(\eps).
 \end{equation}
   Combined with Equation~\ref{eq:z=verM} and Lemma~\ref{lem:computational-triangle0}, and note that $\norm{(\Rej,\bot)-(d,z_{ideal})}=0$, we have that
   \begin{equation}
       \norm{(d,z_{naive})-(d,z_{ideal})}_q\leq O(\eps)
   \end{equation}
\end{proof}

@@

\hannote{intuition about proof}

\hannote{Intuitions about the partition lemma. efficient projectors into passing testing part and not passing testing part.}

\hannote{related notation back to 3.5}

%Let $S_m$ be sets of $\{c\}$ such that only one of the $c_i=1$.
\begin{lemma}[partition lemma \hannote{ref}\Ethan{Lemma number?}]\label{lem:partition2}
Let $(U_0,U)$ be a prover's strategy in Protocol~\ref{proto:QPIP0samp}, where $U_0$ is the how the prover generates answer in the second round, and $U$ is how the prover generates answer in the fourth round. Denote the string $0^{i-1}10^{m-i} \in \zo^m $ as $e_i$.  Let $\gamma_0 \in[0,1]$, and $T\in \mathbb{N}$ such that $\gamma_0=1/\poly(n)$ and $T=\poly(n)$.

@@

By Lemma~\ref{lem:trivial-4-round-strategy} and property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}, the above strategy is accepted with probability $1-\negl(n)$ when $c=0$. Therefore by Lemma~\ref{lem:naive-qpip0-binding}, 
\begin{align}
    \norm{(d,z_{composite,i})-(d,z_{ideal,i})}_q \leq O(\eps)
\end{align}

>>>>

In this section, we combine \cref{QPIP1thm} with results from \cite{FOCS:Mahadev18a} to create a delegation protocol for $\SampBQP$ for fully classical clients. To boost the soundness, we need  polynomial \Ethan{Wording is strange/misleading} rounds, so we use results from \cite{parallelrep} to run polynomially many copies in parallel and get a constant-round protocol.

@@

\subsection{Mahadev's measurement protocol}

@@

\Ethan{Write a bit of technical intro. allows prover commit a state and verifier chooses h allows verifier to learn... MXZ introduce here}

\Ethan{Discussion: Intuitively the purpose of this protocol is to let the verifier learn a measurement outcome. Define MXZ rho h notation here.}

\Ethan{Check tone, this is intro/interface}

\begin{protocol}{Mahadev's measurement protocol $\PiMeasure=(\PMeasure, \VMeasure)$}
\label{proto:urmila4}
\Ethan{Still rooms for improvements, but it's not as bad as before}

Inputs:
\begin{itemize}
	\item Common input: Security parameter $1^\lambda$ where $\lambda\in\bbN$.
	\item Prover's input: a state $\rho\in\cB^{\otimes n}$ for the verifier to measure.
	\item Verifier's inputs:
		\begin{itemize}
			\item a string $h \in \{0,1\}^n$ that specifies the measurement he wants to make: if $h_i=0$, he wants a $X$ measurement on $i$-th qubit. If $h_i=1$, he wants a $Z$ measurement on $i$-th qubit.  
			\item a challenge bit $c\in\zo$. The prover needs to do a ``Testing round" if $c=0$. The prover needs to do a ``Hadamard round" if $c=1$. \Ethan{TODO rephrase. c indicates testing vs Hadamard...}
			\end{itemize}
\end{itemize}

The protocol: \Ethan{Bring back V1 P1 notations}
\begin{enumerate}
    \item The verifier generates a public and secret key pair $(pk, sk)$ on inputs $h$ and $1^\lambda$. It sends $pk$ to the prover.
    \item The prover generates a classical ``commitment" $y$ and internal state $\sigma$ on inputs $pk$ and $\rho$. He send $y$ to the verifier.
	\item The verifier sends $c$ to the prover.
    \item The prover generates a classical string $a$ on inputs $c$ and $\sigma$, and sends it to back to the verifier.
    \item \Ethan{Need more discussion about what's testing/Hadamard} If $c=0$, the verifier \Ethan{performs testing and} decides whether he is going to reject on inputs $(pk,y,a)$.
	If $c=1$, the verifier \Ethan{performs Hadamard round and generates} computes an $n$-bit string $z=V_f(sk,h,y,a)$ on inputs $(sk,h,y,a)$.

	\Ethan{Switch to the following wording:

	If it's a testing round (c=0), then the verifier...
	If it's a Hadamard round (c=1), then the verifier...}
\end{enumerate}
\end{protocol}

\Ethan{Discussion: Notice that testing round has no outputs, then binding property. Then fact.}

Two modification have been made from the original protocol.
\begin{itemize}
	\item On the protocol's third message. Instead of randomly picking whether to run what's called a ``testing round" or a ``Hadamard round", we let the verifier decide as one of its inputs. \Ethan{Remove this and make $c$ random. Instead of this, say that we do parallel repetition but we change the distribution of $c$}
	\item On the protocol's fourth message, the prover can choose to abort by sending $a=\bot$. This is useless for the cheating prover but useful for proving one of our theorems. \Ethan{Move to later when it's used}
\end{itemize}

% \begin{definition}
% With loss of generality,  we can assume that the testing round verification is deterministic. We denote the set of strings $a$ \Ethan{Font makes this $a$ look screwy} accepted by the verifier in the testing round with $(pk,y)$ as $\Acc_{pk,y}$. \hannote{do  i actually use this}
% \end{definition}

\begin{lem}[\cite{FOCS:Mahadev18a}\hannote{change to fact?} \Ethan{Ok I'll do that later}]~\label{lem:trivial-4-round-strategy}
    For any input, there exist a prover strategy $(U_{trivial,0},U_{trivial})$ \Ethan{Consider delay introducing the particular notation; just write this name when it's used later} for Protocol~\ref{proto:urmila4} that is accepted with probability $1-\negl{\lambda}$ \Ethan{Changed n to lambda here, see what does it affect} in the testing.
\end{lem}

@@

\begin{lemma}\label{lem:urmila-binding}[binding property of Protocol~\ref{proto:urmila4}\cite{FOCS:Mahadev18a}]
If a prover's strategy in Protocol~\ref{proto:urmila4}\hannote{need clarification} results in passing the testing round with probability $1-\negl(n)$, then there must exist a $\rho$ such that for all $h$, the verifier will produce a $z$ on the Hadamard round that is $\negl(n)$-Q-computationally \Ethan{Remove Q} indistinguishable\footnote{Recall that quantum computational indistinguishability is defined in Definition~\ref{def:comp-indis}.} from $M_{XZ}(\rho,h)$\hannote{add notation}, unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

@@

By combining the commitment Protocol, $\PiMeasure$, and our $\QPIP_1$ Protocol for $\SampBQP$, $\PiSamp=(\PSamp, \VSamp)$, we get a naive $\QPIP_0$ Protocol for $\SampBQP$ that has no soundness guarantee.
\Ethan{The following discussion is mine}
In particular, we combine them by replacing \Ethan{Still a bit ambiguous. More discussion than just ``replace". Mention it's QPIP1 and whatnot.} the verifier's measurements with a call to $\PiMeasure$. It is possible because $\VSamp$ applies only $X$ and $Z$ measurements non-adaptively.

\begin{protocol}{Intermediate $\QPIP_0$ protocol $\PiNaive$ for a $\SampBQP$ problem $(D_x)_{x\in\set{0, 1}^*}$}
	\label{proto:qpip0_naive}

	Inputs:
	\begin{itemize}
		\item Security parameter $1^\lambda$ where $\lambda\in\bbN$
		\item Error parameter $\varepsilon\in(0, 1)$
		\item Classical input $x\in\zo^n$ to the $\SampBQP$ instance
	\end{itemize}

@@

	Protocol:
	\begin{enumerate}
		\item The verifier choose a $XZ$-measurement $h$ from the distribution specified in Step~\ref{step:qpip1-verify} of $\PiSamp$. \Ethan{I didn't change the wording here}
		It samples $c\xleftarrow{\$}\zo$ uniformly at random.
		\item The prover prepares $\sigma$ by running Step~\ref{step:qpip1-state-gen} of $\PiSamp$.
		\item \label{step:urmila-in-naive} The verifier and prover run $(\PMeasure(\sigma), \VMeasure(h))(1^\lambda, 1^{1/\varepsilon}, x)$.
		\Ethan{Let z be the verifier output when it's a Hadamard round}
		If $c=1$, the verifier obtains $z_{commit}$ \Ethan{measure/commit... so many notations. Confusing. Just do v}.
		\item If $c=1$, the verifier finishes the verification step of Protocol~\ref{ProtoQPIP1} by using $z_{commit}$ as the measurement result he would get and outputs $(d,z_{naive})$ \Ethan{Just do z here}.
	\end{enumerate}
\end{protocol}

\Ethan{Try to update notation:
Use something else for measurement outcome for proto 3. Maybe v or w.
Let z be sampbqp results.}

@@

\begin{equation} \label{eq:z=verM}
    \norm{(d,z_{naive})-(Ver( M_{XZ}(\rho,h))+p_{abort}(\Rej,\bot))}_q \leq \negl(n).
\end{equation}
\hannote{actually true with subnormalize shit?}
 By Theorem~\ref{QPIP1thm}, for all $\rho$, 
 \begin{equation}
     \norm{Ver(M_{XZ}(\rho,h))-(d,z_{ideal})} \leq O(\eps).
 \end{equation}
   Combined with Equation~\ref{eq:z=verM} and Lemma~\ref{lem:computational-triangle0}, and note that $\norm{(\Rej,\bot)-(d,z_{ideal})}=0$, we have that
   \begin{equation}
       \norm{(d,z_{naive})-(d,z_{ideal})}_q\leq O(\eps)
   \end{equation}
\end{proof}

@@

The intuition behind this theorem is that the we send out $m$ independent copies of the naive $\QPIP_0$ and do testing on randomly chosen $m-1$ copies of them, and do Hadamard round and calculate results from the remaining one copy. Therefore if the prover want to cheat, i.e. send something not binding in the Hadamard round copy, he would be caught with probability $1-1/m$. However, the prover is quantum, so he might create quantum correlations that let him cheat with probability more than $1/m$. To prove that such kind of cheating is impossible, we use the following partition lemma from \cite{parallelrep}, which intuitively says that there exist efficient projectors $G_{0,i}$ and $G_{1,i}$ which splits a quantum state in the prover's   internal space into a part that is rejected with high probability in the testing round of $i$-th copy and a part that is accepted with high probability.  $G_{0,i}$ and $G_{1,i}$ need to be efficient because Protocol~\ref{proto:urmila4} only gives computational indistinguishability. Once we have $G_{0,i}$ and $G_{1,i}$, we can split up the prover's internal state $\ket{\psi}$ like conditional probabilistic events. For example, $G_{0,3} G_{1,5} \ket{\psi}$ corresponds to the events of passing the testing round on fifth copy then fails the the testing round on the third copy. We split the state sub-normalized states corresponding to the prover cheating on different copies and show that the sum of probabilities corresponding to bad states can't be more than $1/m$. {The actual proof is more complicated than this sketch because the $G$'s don't commute with each other, so we need to handle the splitting carefully.}

%Let $S_m$ be sets of $\{c\}$ such that only one of the $c_i=1$.
\begin{lemma}[partition lemma (revision of Lemma 3.5 of \cite{parallelrep}\footnote{$G_{0}$ and $G_{1}$ of this version are created from doing $G$ of \cite{parallelrep} and post-select on the $th$ register being $0$ or $1$ then discard $ph,th,in$. Property~\ref{property:partition-err} corresponds to Property~1. Property~\ref{property:partition-testing}corresponds to Property~4, with $2^{m-1}$ changes to $m-1$ because we only have $m$ possible choices of $\{c\}$. Property~\ref{property:partition-binding} corresponds to Property~5. Property~\ref{property-partition-norm-sum} comes from the fact that $G_0$ and $G_1$ are post-selections of orthogonal results of the same $G$.})]\label{lem:partition2}
Let $(U_0,U)$ be a prover's strategy in Protocol~\ref{proto:QPIP0samp}, where $U_0$ is the how the prover generates answer in the second round, and $U$ is how the prover generates answer in the fourth round. Denote the string $0^{i-1}10^{m-i} \in \zo^m $ as $e_i$.  Let $\gamma_0 \in[0,1]$, and $T\in \mathbb{N}$ such that $\gamma_0=1/\poly(n)$ and $T=\poly(n)$.

@@

By Lemma~\ref{lem:trivial-4-round-strategy} and property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}, the above strategy is accepted with probability $1-\negl(n)$ when $c=0$. Since the prover's strartegy is also efficient, by Lemma~\ref{lem:naive-qpip0-binding}, 
\begin{align}
    \norm{(d,z_{composite,i})-(d,z_{ideal,i})}_q \leq O(\eps)
\end{align}
